---
title: "Analysis Template"
author: "Thorley, J.L., Irvine, R.L., Dalgarno, S.I.J. & Amies-Galonski, E."
bibliography: bibliography.bib
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, comment = NA, results = "asis", cache = FALSE)

source("header.R")
library(poisreport)

nheaders <- 2L

rename <- c(
  "BT" = "Bull Trout",
  "RB" = "Rainbow Trout",
  "map" = "Maps" # cos currently using map causes rendering problems!
)

sort <- c(
  "description", "coef", "glance", "map",
  "RB", "BT"
)

drop <- NULL
```

```{r}
cat(report_draft())
```

The suggested citation for this [analytic appendix](http://www.poissonconsulting.ca/analytic-appendices.html) is:

```{r}
cat(report_citation(file_name()))
```

## Background

Someone has been collecting data since.

The primary goal of the current analyses is to answer the following questions:

> Is y affected by x?

### Data Preparation

The data were proved by someone as an Excel spreadsheet.
The data were prepared for analysis using `r substr(R.version.string,1,15)` [@r_core_team_r:_2019].

Key assumptions of the data preparation included:

- The data were recorded correctly.

### Statistical Analysis

Model parameters were estimated using Maximum-Likelihood (ML) and Bayesian methods.
The Maximum-Likelihood Estimates (MLE) were obtained using TMB [@kristensen_tmb_2016] while the Bayesian estimates were produced using JAGS [@plummer_jags_2015] and STAN [@carpenter_stan_2017].
For additional information on ML and Bayesian estimation the reader is referred to @millar_maximum_2011 and @mcelreath_statistical_2016, respectively.
ML models have the advantage of being relatively quick to fit while Bayesian models capture all the uncertainty and allow intuitive probabilistic statements.

Unless stated otherwise, the Bayesian analyses used weakly informative normal and half-normal prior distributions [@gelman_prior_2017].
The posterior distributions were estimated from `r getOption("mb.nchains") * getOption("mb.niters") ` Markov Chain Monte Carlo (MCMC) samples thinned from the second halves of `r getOption("mb.nchains")` chains [@kery_bayesian_2011, pp. 38-40].
Model convergence was confirmed by ensuring that the potential scale reduction factor $\hat{R} \leq `r round(getOption("mb.rhat"), 2)`$ [@kery_bayesian_2011, pp. 40] and the effective sample size [@brooks_handbook_2011] $\textrm{ESS} \geq  `r getOption("mb.nchains") * getOption("mb.niters") * round(getOption("mb.esr"), 2)`$ for each of the monitored parameters [@kery_bayesian_2011, pp. 61].

The parameters are summarised in terms of the point *estimate*, standard deviation (*sd*), the *z-score*, *lower* and *upper* 95% confidence/credible limits (CLs) and the *p-value* [@kery_bayesian_2011, pp. 37, 42].
For ML models, the point estimate is the MLE, the standard deviation is the standard error, the z-score is $\mathrm{MLE}/\mathrm{sd}$ and the 95% CLs are the $\mathrm{MLE} \pm 1.96 \cdot \mathrm{sd}$.
For Bayesian models, the estimate is the median (50th percentile) of the MCMC samples, the z-score is $\mathrm{mean}/\mathrm{sd}$ and the 95% CLs are the 2.5th and 97.5th percentiles.
A p-value of 0.05 indicates that the lower or upper 95% CL is 0.

Model averaging and selection was implemented using an information theoretic approach [@burnham_model_2002].
ML models were compared using the marginal Akaike's Information Criterion corrected for small sample size [AICc, @burnham_model_2002] while Bayesian models were compared using the Watanabe-Akaike Information Criterion [WAIC, @watanabe_asymptotic_2010].
The support for fixed terms in models with identical *random* [@kery_bayesian_2011, pp. 75] effects was assessed using the Akaike's weights ($w_i$) which were also used for model averaging [@burnham_model_2002]. 
The top ML model for each analysis was refitted as a Bayesian model. 
Where relevant, model adequacy was confirmed by examination of residual plots for the full model(s). 

The results are displayed graphically by plotting the modeled relationships between particular variables and the response(s) with the remaining variables held constant.
In general, continuous and discrete fixed variables are held constant at their mean and first level values, respectively, while random variables are held constant at their typical values (expected values of the underlying hyperdistributions) [@kery_bayesian_2011, pp. 77-82].
When informative the influence of particular variables is expressed in terms of the *effect size* (i.e., percent change in the response variable) with 95% confidence/credible intervals [CIs, @bradford_using_2005].

The analyses were implemented using `r substr(R.version.string,1,15)` [@r_core_team_r:_2019] and the [`mbr`](http://www.poissonconsulting.ca/mbr) family of packages.

### Model Descriptions

#### Count

The count data were analysed using an overdispersed Poisson model.
Key assumptions of the count model include:

- The counts are gamma-Poisson distributed.

### Model Templates

```{r}
cat(sbr_blocks(sort = sort, rename = rename, nheaders = nheaders, drop = drop))
```

## Results

### Tables

```{r}
cat(sbr_tables(sort = sort, rename = rename, nheaders = nheaders, drop = drop))
```

### Figures

```{r}
cat(sbr_figures(sort = sort, rename = rename, nheaders = nheaders, drop = drop))
```

## Acknowledgements

The organisations and individuals whose contributions have made this analytic appendix possible include:

- A Company
    - FirstName LastName
    - FirstName LastName

## References
