---
title: "Llgaay gwii 2017"
author: "Irvine, R.L. & Thorley, J.L."
bibliography: bibliography.bib
---
  
```{r, echo = FALSE, warning = FALSE, message = FALSE, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, comment = NA, results = "asis", cache = FALSE)

source("header.R")
library(poisreport)

nheaders <- 2L

rename <- c(
  "rate" = "Removal",
  "map" = "Maps" # to avoid rendering problems!
)

sort <- c(
  "description", "coef", "glance", "map"
)

drop <- NULL
```

```{r}
cat(report_draft())
```

The suggested citation for this [analytic appendix](https://www.poissonconsulting.ca/analytic-appendices.html) is:
  
```{r}
cat(report_citation(file_name()))
```

## Background

To restore balance, a deer removal program was implemented on various islands in Gwaii Haanas.

The primary goal of the current analyses is to answer the following questions:
  
> What was the relative efficiency of the removal methods?

> How many deer were remaining on the islands?

> What would the cost of removing the remaining deer been?

### Data Preparation
  
The data were provided by Parks Canada in the form of Excel spreadsheets and prepared for analysis using `r substr(R.version.string,1,15)` [@r_core_team_r_2020].

### Statistical Analysis

Model parameters were estimated using Bayesian methods.
The estimates were produced using JAGS [@plummer_jags:_2003].
For additional information on Bayesian estimation the reader is referred to @mcelreath_statistical_2016.

Unless stated otherwise, the Bayesian analyses used weakly informative normal and half-normal prior distributions [@gelman_prior_2017].
The posterior distributions were estimated from `r getOption("mb.nchains", 3L) * getOption("mb.niters", 500L) ` Markov Chain Monte Carlo (MCMC) samples thinned from the second halves of `r getOption("mb.nchains", 3L)` chains [@kery_bayesian_2011, pp. 38-40].
Model convergence was confirmed by ensuring that the potential scale reduction factor $\hat{R} \leq `r round(getOption("mb.rhat", 1.05), 2)`$ [@kery_bayesian_2011, pp. 40] and the effective sample size [@brooks_handbook_2011] $\textrm{ESS} \geq  `r getOption("mb.nchains", 3L) * getOption("mb.niters", 500) * round(getOption("mb.esr", 0.1), 2)`$ for each of the monitored parameters [@kery_bayesian_2011, pp. 61].

The parameters are summarised in terms of the point *estimate*, standard deviation (*sd*), the *z-score*, *lower* and *upper* 95% confidence/credible limits (CLs) and the *p-value* [@kery_bayesian_2011, pp. 37, 42].
The estimate is the median (50th percentile) of the MCMC samples, the z-score is $\mathrm{mean}/\mathrm{sd}$ and the 95% CLs are the 2.5th and 97.5th percentiles.
A p-value of 0.05 indicates that the lower or upper 95% CL is 0.

The results are displayed graphically by plotting the modeled relationships between particular variables and the response(s) with the remaining variables held constant.
In general, continuous and discrete fixed variables are held constant at their mean and first level values, respectively, while random variables are held constant at their typical values (expected values of the underlying hyperdistributions) [@kery_bayesian_2011, pp. 77-82].
When informative the influence of particular variables is expressed in terms of the *effect size* (i.e., percent change or n-fold in the response variable) with 95% confidence/credible intervals [CIs, @bradford_using_2005].

The analyses were implemented using `r substr(R.version.string,1,15)` [@r_core_team_r_2020] and the [`mbr`](https://www.poissonconsulting.ca/mbr) family of packages.

### Model Descriptions

#### Model

The data were analysed using a removal model.

Key assumptions of the model include:

- There is no migration among the islands during the course of the study.
- The efficiency of bait station, helicopter and boat hunting is directly proportional to the deer density.
- The efficiency of all other methods is independent of deer density.
- The effort is the relative cost of the hunting event.
- The expected number of deer removed is the product of the efficiency and effort.
- The actual number of deer removed is gamma-Poisson distributed.

### Model Templates

```{r}
cat(sbr_blocks(sort = sort, rename = rename, nheaders = nheaders, drop = drop))
```

## Results

### Tables

```{r}
cat(sbr_tables(sort = sort, rename = rename, nheaders = nheaders, drop = drop))
```

### Figures

```{r}
cat(sbr_figures(sort = sort, rename = rename, nheaders = nheaders, drop = drop))
```

## Acknowledgements

The organisations and individuals whose contributions have made this analytic appendix possible include:
  
- Parks Canada
  - Nadine
  - Christine Bentley
  - Charlotte
- Poisson Consulting
  - Sebastian Dalgarno

## References
