---
title: "llgaay-gwii 2020"
author: "Irvine, R.L., Thorley, J.L. & Dalgarno, S.I.J."
bibliography: bibliography.bib
---
  
```{r, echo = FALSE, warning = FALSE, message = FALSE, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, comment = NA, results = "asis", cache = FALSE)

source("header.R")
library(poisreport)

nheaders <- 2L

rename <- c(
  "map" = "Maps" # to avoid rendering problems!
)

sort <- c(
  "description", "coef", "glance", "map"
)

drop <- NULL
```

```{r}
cat(report_draft())
```

The suggested citation for this [analytic appendix](https://www.poissonconsulting.ca/analytic-appendices.html) is:
  
  ```{r}
cat(report_citation(file_name()))
```

## Background

The primary goal of the current analyses is to answer the following questions:
  
  > Is y affected by x?
  
### Data Preparation
  
The data were prepared for analysis using `r substr(R.version.string,1,15)` [@r_core_team_r_2020].

Key assumptions of the data preparation included:
  
- The data were recorded correctly.

### Statistical Analysis

Model parameters were estimated using Bayesian methods.
The estimates were produced using JAGS [@plummer_jags:_2003] and STAN [@carpenter_stan_2017].
For additional information on Bayesian estimation the reader is referred to @mcelreath_statistical_2016, respectively.

Unless stated otherwise, the Bayesian analyses used weakly informative normal and half-normal prior distributions [@gelman_prior_2017].
The posterior distributions were estimated from `r getOption("mb.nchains", 3L) * getOption("mb.niters", 500L) ` Markov Chain Monte Carlo (MCMC) samples thinned from the second halves of `r getOption("mb.nchains", 3L)` chains [@kery_bayesian_2011, pp. 38-40].
Model convergence was confirmed by ensuring that the potential scale reduction factor $\hat{R} \leq `r round(getOption("mb.rhat", 1.05), 2)`$ [@kery_bayesian_2011, pp. 40] and the effective sample size [@brooks_handbook_2011] $\textrm{ESS} \geq  `r getOption("mb.nchains", 3L) * getOption("mb.niters", 500) * round(getOption("mb.esr", 0.1), 2)`$ for each of the monitored parameters [@kery_bayesian_2011, pp. 61].

The parameters are summarised in terms of the point *estimate*, standard deviation (*sd*), the *z-score*, *lower* and *upper* 95% confidence/credible limits (CLs) and the *p-value* [@kery_bayesian_2011, pp. 37, 42].
The estimate is the median (50th percentile) of the MCMC samples, the z-score is $\mathrm{mean}/\mathrm{sd}$ and the 95% CLs are the 2.5th and 97.5th percentiles.
A p-value of 0.05 indicates that the lower or upper 95% CL is 0.

The results are displayed graphically by plotting the modeled relationships between particular variables and the response(s) with the remaining variables held constant.
In general, continuous and discrete fixed variables are held constant at their mean and first level values, respectively, while random variables are held constant at their typical values (expected values of the underlying hyperdistributions) [@kery_bayesian_2011, pp. 77-82].
When informative the influence of particular variables is expressed in terms of the *effect size* (i.e., percent change or n-fold in the response variable) with 95% confidence/credible intervals [CIs, @bradford_using_2005].

The analyses were implemented using `r substr(R.version.string,1,15)` [@r_core_team_r_2020] and the [`mbr`](https://www.poissonconsulting.ca/mbr) family of packages.

### Model Descriptions

#### Model

The data were analysed using a model.
Key assumptions of the model include:
  
- The data are normally distributed.

### Model Templates

```{r}
cat(sbr_blocks(sort = sort, rename = rename, nheaders = nheaders, drop = drop))
```

## Results

### Tables

```{r}
cat(sbr_tables(sort = sort, rename = rename, nheaders = nheaders, drop = drop))
```

### Figures

```{r}
cat(sbr_figures(sort = sort, rename = rename, nheaders = nheaders, drop = drop))
```

## Acknowledgements

The organisations and individuals whose contributions have made this analytic appendix possible include:
  
- A Company
  - FirstName LastName
- FirstName LastName

## References
